---
title: "METABO -- QC for one sample"
author: "ATIA Safiya"
date: "2023-07-26"
params:
  samp: "samp"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
---

# Seurat objects  
## Load required packages  
```{r,  message=FALSE, warning=FALSE}
library(Seurat)
library(tidyverse)
library(ggplot2)
library("scales")
```

## Load seurat object (ALL GENES)
```{r}
samp <- params$samp
```

```{r}
seurat_obj <- readRDS(file = paste0('./Lucille/results/rds/samples/Norm_', samp,'.rds'))

seurat_obj
table(seurat_obj$orig.ident)
```

```{r}
counts <- GetAssayData(seurat_obj, assay = "RNA")
print(length(rownames(counts)))

counts <- GetAssayData(seurat_obj, assay = "SCT")
print(length(rownames(counts)))
```

## Create seurat object (METABO)
```{r}
genes <- read.csv("all_metabo.tsv", sep = '\t', header = TRUE)
head(genes, 3)
```

```{r}
features <- rownames(seurat_obj@assays$SCT@counts)
# keep genes names seeen un genes$Gene_name
matched_features <- features[features %in% genes$Gene_name]

counts <- GetAssayData(seurat_obj, assay = "SCT")
counts <- counts[(which(rownames(counts) %in% matched_features)),]
metabo_obj <- subset(seurat_obj, features = rownames(counts))
```

```{r}
metabo_obj
table(metabo_obj$orig.ident)
```

```{r}
counts <- GetAssayData(metabo_obj, assay = "RNA")
print(length(rownames(counts)))

counts <- GetAssayData(metabo_obj, assay = "SCT")
print(length(rownames(counts)))
```

##  <span style="color:#296d98;"> Parameters </span>  
```{r}
pc = 60
```

rerun SCTransform after subsetting the object, especially if you intend to rerun PCA (i.e. subclustering)

# QC metrics  

```{r, message=FALSE, warning=FALSE}
metadata <- metabo_obj@meta.data
```

## Nb of reads  
```{r, message=FALSE}
VlnPlot(metabo_obj,features="nCount_RNA", group.by = "orig.ident", pt.size = 0)
RidgePlot(metabo_obj, features = "nCount_RNA", group.by = "orig.ident")

metadata %>% 
  ggplot(aes( x=nCount_RNA,y=orig.ident, fill= orig.ident)) + 
	geom_boxplot() + 
	theme_classic()+
  scale_x_log10()
```
```{r}
nCount_metabo <- as.data.frame(rowSums(metabo_obj[["nCount_RNA"]]))
nCount_metabo$Genes <- "metabo"
colnames(nCount_metabo)[1] <- "nCount"

nCount_all <- as.data.frame(rowSums(seurat_obj[["nCount_RNA"]]))
nCount_all$Genes <- "all"
colnames(nCount_all)[1] <- "nCount"

other <- nCount_all
other$Genes <- "others"
dim(other)
other <- other[!(rownames(other) %in% rownames(metabo_obj$SCT)),]

expr <- rbind(nCount_all, nCount_metabo, other)
head(expr)
ggplot(expr, aes(x=Genes,y=nCount)) + geom_boxplot()
```

## Nb of genes  
```{r, message=FALSE}
VlnPlot(metabo_obj,features="nFeature_RNA", group.by = "orig.ident", pt.size = 0)
RidgePlot(metabo_obj, features = "nFeature_RNA", group.by = "orig.ident")

metadata %>% 
  ggplot(aes( x=nFeature_RNA,y=orig.ident, fill= orig.ident)) + 
	geom_boxplot() + 
	theme_classic()+
    scale_x_log10()
    scale_x_log10()
    
  scale_x_log10()
    
```

```{r}
ggplot(metabo_obj[[]]) + 
   geom_histogram(aes(nFeature_RNA), 
                  color = "#558bdc", fill= "#173664",
                  binwidth = 50) + 
  ggtitle("Distribution of nFeature_RNA ") + NoLegend()
```

## Gene expression  
We can pick the first 100 cells and look at the distributions of their expression values.

```{r, warning=FALSE}
as.tibble(
  metabo_obj@assays$SCT@data[,1:100]
) %>%
  pivot_longer(
    cols=everything(),
    names_to="cell",
    values_to="expression"
  ) %>%
  ggplot(aes(x=expression, group=cell)) +
  geom_density() +
  coord_cartesian(ylim=c(0,0.6), xlim=c(0,3))
```

```{r}
expression_metabo <- as.data.frame(rowSums(metabo_obj@assays$SCT@data))
expression_metabo$Genes <- "metabo"
colnames(expression_metabo)[1] <- "expression"

expression_all <- as.data.frame(rowSums(seurat_obj@assays$SCT@data))
expression_all$Genes <- "all"
colnames(expression_all)[1] <- "expression"

other <- expression_all
other$Genes <- "others"
dim(other)
other <- other[!(rownames(other) %in% rownames(metabo_obj$SCT)),]

expr <- rbind(expression_all, expression_metabo, other)
head(expr)
ggplot(expr, aes(x=Genes,y=log(expression+1))) + geom_boxplot()
```

## MT genes  
### All genes
```{r, message=FALSE}
VlnPlot(metabo_obj, features = "percent.mt", group.by = "orig.ident", pt.size = 0)
RidgePlot(metabo_obj, features = "percent.mt", group.by = "orig.ident")
```

```{r}
ggplot(metabo_obj[[]]) + 
   geom_histogram(aes(percent.mt), 
                  color = "#558bdc", fill= "#173664",
                  binwidth = 0.5) + 
  ggtitle("Distribution of Percentage Mitochondrion") + NoLegend()
```

### Metabo only
```{r, message=FALSE}
metabo_obj[["percent.mt"]] <- PercentageFeatureSet(metabo_obj, pattern = "^MT-", assay = 'RNA')
VlnPlot(metabo_obj, features = "percent.mt", group.by = "orig.ident", pt.size = 0)
RidgePlot(metabo_obj, features = "percent.mt", group.by = "orig.ident")
```

```{r}
ggplot(metabo_obj[[]]) + 
   geom_histogram(aes(percent.mt), 
                  color = "#558bdc", fill= "#173664",
                  binwidth = 0.5) + 
  ggtitle("Distribution of Percentage Mitochondrion") + NoLegend()
```

## Ribosomal genes  
### All genes  
```{r, message=FALSE}
VlnPlot(metabo_obj, features = "percent.ribosomal", group.by = "orig.ident", pt.size = 0)
RidgePlot(metabo_obj, features = "percent.ribosomal", group.by = "orig.ident")
```

```{r}
ggplot(metabo_obj[[]]) + 
   geom_histogram(aes(percent.ribosomal), 
                  color = "#558bdc", fill= "#173664",
                  binwidth = 0.7) + 
  ggtitle("Distribution of Percentage Ribosomal") + NoLegend()
```

### Metabo only  
```{r, message=FALSE}
PercentageFeatureSet(metabo_obj,pattern="^RP[LS]", assay = 'RNA') -> metabo_obj[["percent.ribosomal"]] 
VlnPlot(metabo_obj, features = "percent.ribosomal", group.by = "orig.ident", pt.size = 0)
RidgePlot(metabo_obj, features = "percent.ribosomal", group.by = "orig.ident")
```

```{r}
ggplot(metabo_obj[[]]) + 
   geom_histogram(aes(percent.ribosomal), 
                  color = "#558bdc", fill= "#173664",
                  binwidth = 0.7) + 
  ggtitle("Distribution of Percentage Ribosomal") + NoLegend()
```

## Largest gene  (All genes)
```{r, message=FALSE}
VlnPlot(metabo_obj, features = "percent.largest_gene", group.by = "orig.ident", pt.size = 0)
RidgePlot(metabo_obj, features = "percent.largest_gene", group.by = "orig.ident")
```

```{r}
ggplot(metabo_obj[[]]) + 
   geom_histogram(aes(percent.largest_gene), 
                  color = "#558bdc", fill= "#173664",
                  binwidth = 0.7) + 
  ggtitle("Distribution of Percentage Largest Gene") + NoLegend()
```

## FeatureScatter  
```{r}
FeatureScatter(metabo_obj, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by ="orig.ident")

FeatureScatter(metabo_obj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA",group.by ="orig.ident") + geom_smooth(method = 'lm')
```

## Cell cycle scoring  
```{r, warning=FALSE}
metabo_obj@meta.data %>%
  group_by(orig.ident,Phase) %>%
  count() %>%
  group_by(orig.ident) %>%
  mutate(percent=100*n/sum(n)) %>%
  ungroup() %>%
  ggplot(aes(x=orig.ident,y=percent, fill=Phase)) +
  scale_fill_manual( values=c(G1='azure4',G2M='dodgerblue',S='tomato3' )) +
  geom_col() +
  ggtitle("Percentage of cell cycle phases per sample")

as_tibble(metabo_obj[[]]) %>%
  ggplot(aes(Phase, fill=Phase)) +
  scale_fill_manual( values=c(G1='azure4',G2M='dodgerblue',S='tomato3' ))+ geom_bar()


as_tibble(metabo_obj[[]]) %>%
  ggplot(aes(x=S.Score, y=G2M.Score, color=Phase)) + 
  scale_color_manual( values=c(G1='azure4',G2M='dodgerblue',S='tomato3' )) +
  geom_point() +
  coord_cartesian(xlim=c(-0.15,0.15), ylim=c(-0.15,0.15))
```

# Identify high variable features  
*Included in the `SCTransform` function. It was done with all genes*

```{r, message=FALSE, warning=FALSE}
metabo_obj <- FindVariableFeatures(metabo_obj, selection.method = "vst", nfeatures = 1000)

# identify the 10 most highly variable genes
top10 <- head(VariableFeatures(metabo_obj), 10)
top10

# plot the variable features 
plot <- VariableFeaturePlot(metabo_obj)
LabelPoints(plot = plot, points = head(top10,10), repel = TRUE)
```

We can plot out a graph of the variance vs mean and highlight the selected genes so we can see whether we think we’re likely to capture what we need.    

```{r, eval=FALSE}
as_tibble(HVFInfo(metabo_obj),rownames = "Gene") -> variance.data

variance.data %>% 
  mutate(hypervariable=Gene %in% VariableFeatures(metabo_obj)
) -> variance.data

variance.data %>% 
  ggplot(aes(log(mean),log(variance),color=hypervariable)) + 
  geom_point() + 
  scale_color_manual(values=c("black","red"))
```
  

# Scaling the data  
*Included in the `SCTransform` function. It was done with all genes*

**ScaleData** :

* Shifts the expression of each gene, so that the mean expression across cells is 0
* Scales the expression of each gene, so that the variance across cells is 1 
    + This step gives equal weight in downstream analyses, so that highly-expressed genes do not 

# Linear dimensional reduction  
## PCA  
```{r}
DefaultAssay(metabo_obj) <- "SCT"

metabo_obj <- RunPCA(metabo_obj, verbose=FALSE)
VizDimLoadings(metabo_obj, dims = 1:2, reduction = "pca")
```

```{r}
mat <- Seurat::GetAssayData(metabo_obj, assay = "SCT", slot = "scale.data")
pca <- metabo_obj[["pca"]]

# Get the total variance:
total_variance <- sum(matrixStats::rowVars(mat))
eigValues = (pca@stdev)^2  ## EigenValues

varExplained = eigValues / total_variance

pc1 = percent(varExplained[1], accuracy = 0.01)
pc2 = percent(varExplained[2], accuracy = 0.01)
```


```{r}
DimPlot(metabo_obj, reduction = "pca", group.by = "orig.ident") + xlab(paste0("PC_1 : ", pc1)) + ylab(paste0("PC_2 :", pc2 ))
DimPlot(metabo_obj, reduction = "pca", group.by = "Phase", cols = c(G1='azure4',G2M='dodgerblue',S='tomato3'))

DimPlot(metabo_obj, reduction = "pca", group.by = "orig.ident", label.size = 4, label = TRUE)
```

### Save seurat object (metabo)
```{r}
saveRDS(metabo_obj, file = paste0("./Lucille/results/rds/samples/", samp, "_metabo.rds"))
```

## Determine the dimensionality of the dataset  
how many components should we choose to include?  

**JackStrawPlot**  
*significant* PCs as those who have a strong enrichment of **low p-value**

* dashed samp : uniform distribution
* solid curve above dash samp : strong enrichments of features with low p-values

*Cannot run Jackstraw plot on SCT data*  
```{r, eval=FALSE}
metabo_obj <- JackStraw(metabo_obj, num.replicate = 100, dims = pc)
jack.data <- metabo_obj 

metabo_obj <- ScoreJackStraw(metabo_obj, dims = 1:pc)

JackStrawPlot(metabo_obj, dims = 1:pc)
```

**ElbowPlot**  
ranking of principle components based on the percentage of variance explained by each one 

```{r, warning=FALSE, message=FALSE}
ElbowPlot(seurat_obj, ndims = pc)
```