---
title: "QC plots unfiltered"
author: "ATIA Safiya"
date: "2023-07-26"
params:
  sample: "sample"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
---

# Environment  
## Load required packages  
```{r,  message=FALSE, warning=FALSE}
library(Seurat)
library(tidyverse)
library(viridis)
```

```{r,  message=FALSE, warning=FALSE}
sample <- params$sample # sample_condition
condition <- str_split_1(sample, '_')[2]
```

## Loading data  

```{r}
seurat_obj <- readRDS(file = paste0("./results/rds/mtx/mtx_", sample, ".rds"))
                
seurat_obj                  
```

# QC metrics (RAW)
**nCount_RNA** : the total number of reads (or more correctly UMIs) in the dataset  
**nFeature_RNA** : the number of observed genes (anything with a nonzero count)

##  <span style="color:#296d98;"> Parameters </span>  
```{r}
## nCount
x1 = 500
## nFeature
x2 = 220
## Mitochondrion
x3 = 15
```

## Nb of reads  
```{r, message=FALSE}
ggplot(seurat_obj[[]]) + 
   geom_histogram(aes(nCount_RNA, 
                  color = ifelse(nCount_RNA < x1 , "a", "b")),
                  binwidth = 50) + 
  scale_color_manual(values = c("a" = "#558bdc", "b" = "#173664")) +
  theme_classic()+
  xlim(c(0, 50000))+
  ylim(c(0, 50))+
  ggtitle("Distribution of nCount_RNA ") + NoLegend()
```

## Nb of genes  
```{r}
ggplot(seurat_obj[[]]) + 
   geom_histogram(aes(nFeature_RNA, 
                  color = ifelse(nFeature_RNA < x2 , "a", "b")),
                  binwidth = 50) + 
  scale_color_manual(values = c("a" = "#558bdc", "b" = "#173664")) +
  xlim(c(0, 12000))+ 
  ylim(c(0, 100))+
  ggtitle("Distribution of nFeature_RNA ") + NoLegend()
```


## MT genes  

```{r}
seurat_obj[["percent.mt"]] <- PercentageFeatureSet(seurat_obj, pattern = "^MT-")

ggplot(seurat_obj[[]]) + 
   geom_histogram(aes(percent.mt, 
                  color = ifelse(percent.mt < x3 , "a", "b"))) + 
  scale_color_manual(values = c("a" = "#173664", "b" = "#558bdc")) + 
  xlim(c(0, 100))+ 
  ylim(c(0, 100))+
  ggtitle("Distribution of Percentage Mitochondrion") + NoLegend()
```

## Ribosomal genes  

```{r}
PercentageFeatureSet(seurat_obj,pattern="^RP[LS]") -> seurat_obj[["percent.ribosomal"]] 

ggplot(seurat_obj[[]]) + 
   geom_histogram(aes(percent.ribosomal), 
                  color = "#173664", fill= "#558bdc") + 
    xlim(c(0, 30))+ 
  ylim(c(0, 200))+
  ggtitle("Distribution of Percentage Ribosomal") + NoLegend()
```



# Filtering  
```{r}
rm(seurat_obj)
seurat_obj <- readRDS(file = paste0("./results/rds/samples/Filtered_", sample, ".rds"))
                
seurat_obj                  
```

# QC metrics (after filtering)
**nCount_RNA** : the total number of reads (or more correctly UMIs) in the dataset  
**nFeature_RNA** : the number of observed genes (anything with a nonzero count)

## Nb of reads  
```{r, message=FALSE}
ggplot(seurat_obj[[]]) + 
   geom_histogram(aes(nCount_RNA, 
                  color = ifelse(nCount_RNA < x1 , "a", "b")),
                  binwidth = 50) + 
  scale_color_manual(values = c("a" = "#558bdc", "b" = "#173664")) +
  theme_classic()+
  xlim(c(0, 50000))+
  ylim(c(0, 50))+
  ggtitle("Distribution of nCount_RNA ") + NoLegend()
```

## Nb of genes  
```{r}
ggplot(seurat_obj[[]]) + 
   geom_histogram(aes(nFeature_RNA, 
                  color = ifelse(nFeature_RNA < x2 , "a", "b")),
                  binwidth = 50) + 
  scale_color_manual(values = c("a" = "#558bdc", "b" = "#173664")) +
  xlim(c(0, 12000))+ 
  ylim(c(0, 100))+
  ggtitle("Distribution of nFeature_RNA ") + NoLegend()
```


## MT genes  

```{r}
seurat_obj[["percent.mt"]] <- PercentageFeatureSet(seurat_obj, pattern = "^MT-")

ggplot(seurat_obj[[]]) + 
   geom_histogram(aes(percent.mt, 
                  color = ifelse(percent.mt < x3 , "a", "b"))) + 
  scale_color_manual(values = c("a" = "#173664", "b" = "#558bdc")) + 
  xlim(c(0, 100))+ 
  ylim(c(0, 100))+
  ggtitle("Distribution of Percentage Mitochondrion") + NoLegend()
```

## Ribosomal genes  

```{r}
PercentageFeatureSet(seurat_obj,pattern="^RP[LS]") -> seurat_obj[["percent.ribosomal"]] 

ggplot(seurat_obj[[]]) + 
   geom_histogram(aes(percent.ribosomal), 
                  color = "#173664", fill= "#558bdc") + 
    xlim(c(0, 30))+ 
  ylim(c(0, 200))+
  ggtitle("Distribution of Percentage Ribosomal") + NoLegend()
```
