---
title: "Markers of clusters for merged samples [Normalized]"
author: "ATIA Safiya"
date: "2023-07-27"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
---

# Objects 
## Load required packages
```{r,  message=FALSE, warning=FALSE}
library(Seurat)
library(tidyverse)
library(ggplot2)
options(future.globals.maxSize = 8000 * 1024^2)
```

## Load markers

```{r}
all_clusters <- readRDS(file = "./results/html/all_samples_cluster.markers.rds")
dim(all_clusters)
```

```{r}
all_metabo <- readRDS(file = "./results/html/all_samples_cluster.markers_metabo.rds")
dim(all_metabo)
```


```{r}
seurat_obj <- readRDS(file = './results/rds/all_samples_Clusters.rds')

seurat_obj
table(seurat_obj$orig.ident)
```


```{r}
metabo_obj <- readRDS(file = './results/rds/all_samples_Clusters_metabo.rds')

metabo_obj
table(metabo_obj$orig.ident)
```

## Biomarkers all genes  

* `DoHeatmap` generates an expression heatmap for given cells and features. In this case, we are plotting the top 20 markers (or all markers if less than 20) for each cluster.   

```{r}
prolif <- all_clusters[which( all_clusters$diff_p_val == all_clusters$max_pval), ]
diff <- all_clusters[which( all_clusters$prolif_p_val == all_clusters$max_pval), ]
dim(prolif)
dim(diff)
```

```{r}
prolif <- prolif[, c('prolif_p_val', 'prolif_avg_log2FC', 'prolif_p_val_adj', 'gene', 'cluster' )]
head(prolif)
```

```{r}
diff <- diff[, c('diff_p_val', 'diff_avg_log2FC', 'diff_p_val_adj', 'gene', 'cluster' )]
head(diff)
```

```{r}
diff_markers <- diff[order(diff$diff_p_val_adj, decreasing = FALSE),]
prolif_markers <- diff[order(prolif$prolif_p_val_adj, decreasing = FALSE),]
```

```{r, warning=FALSE, message=FALSE}
diff_markers %>%
    group_by(cluster) %>%
    top_n(n = 10, wt = diff_avg_log2FC) -> top10

top10
DoHeatmap(seurat_obj, features = top10$gene) + NoLegend()

prolif_markers %>%
    group_by(cluster) %>%
    top_n(n = 10, wt = diff_avg_log2FC) -> top10

top10
DoHeatmap(seurat_obj, features = top10$gene) + NoLegend()
```


## Biomarkers metabo genes  

* `DoHeatmap` generates an expression heatmap for given cells and features. In this case, we are plotting the top 20 markers (or all markers if less than 20) for each cluster.   

```{r}
prolif <- all_metabo[which( all_metabo$diff_p_val == all_metabo$max_pval), ]
diff <- all_metabo[which( all_metabo$prolif_p_val == all_metabo$max_pval), ]
dim(prolif)
dim(diff)
```

```{r}
prolif <- prolif[, c('prolif_p_val', 'prolif_avg_log2FC', 'prolif_p_val_adj', 'gene', 'cluster' )]
head(prolif)
```

```{r}
diff <- diff[, c('diff_p_val', 'diff_avg_log2FC', 'diff_p_val_adj', 'gene', 'cluster' )]
head(diff)
```

```{r}
diff_markers <- diff[order(diff$diff_p_val_adj, decreasing = FALSE),]
prolif_markers <- diff[order(prolif$prolif_p_val_adj, decreasing = FALSE),]
```

```{r, warning=FALSE, message=FALSE}
diff_markers %>%
    group_by(cluster) %>%
    top_n(n = 10, wt = diff_avg_log2FC) -> top10

top10
DoHeatmap(metabo_obj, features = top10$gene) + NoLegend()

prolif_markers %>%
    group_by(cluster) %>%
    top_n(n = 10, wt = diff_avg_log2FC) -> top10
    
top10
DoHeatmap(metabo_obj, features = top10$gene) + NoLegend()
```