---
title: "Defining clusters for merged samples [Normalized]"
author: "ATIA Safiya"
date: "2023-06-26"
output:
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
---

# Seurat object  
## Load required packages
```{r,  message=FALSE, warning=FALSE}
library(Seurat)
library(tidyverse)
library(ggplot2)
library(clustree)
library("future")
```

## Load seurat object (METABO)
```{r}
metabo_obj <- readRDS(file = "merged_seurat_metabo.rds")

metabo_obj
table(metabo_obj$orig.ident)
```

# Non linear reduction  
##  <span style="color:#296d98;"> Parameters </span>  
```{r}
pc = 30
8482 -> saved.seed
set.seed(saved.seed)
thread = 8
```

## UMAP
```{r, warning=FALSE}
metabo_obj <- RunUMAP(metabo_obj, dims = 1:pc, seed.use = saved.seed, )

DimPlot(metabo_obj, reduction = "umap", group.by = "orig.ident")
DimPlot(metabo_obj, reduction = "umap", group.by = "Phase", cols = c(G1='azure4',G2M='dodgerblue',S='tomato3') )
```

## tSNE
tSNE with default Perplexity 30 :
```{r}
metabo_obj <- RunTSNE(metabo_obj, dims=1:pc,seed.use = saved.seed)

DimPlot(metabo_obj, reduction = "tsne", group.by = "orig.ident")
DimPlot(metabo_obj, reduction = "tsne", group.by = "Phase", cols = c(G1='azure4',G2M='dodgerblue',S='tomato3'))
```

# Defining cell clusters  
`FindNeighbors`:    

* construction of a KNN graph based on the euclidean distance in PCA space, and refine the edge weights between any two cells based on the shared overlap in their local neighborhoods (Jaccard similarity).
* *input* : the previously defined dimensionality of the dataset.

`FindClusters`:   

* default : Louvain algorithm  
* iteratively group cells together, with the goal of optimizing the standard modularity function.  
* *Resolution* : setting this parameter between 0.4-1.2 typically returns good results for single-cell datasets of around 3K cells

```{r}
metabo_obj <- FindNeighbors(metabo_obj, dims = 1:pc)
```

## Save seurat object  
```{r}
saveRDS(metabo_obj, file ="merged_seurat_FindNeighbors_metabo.rds")
```

## Optimize resolution  
```{r, message=FALSE}
plan("multisession", workers = thread)
metabo_obj <- FindClusters(metabo_obj, resolution = 0.2)

plan("multisession", workers = thread)
metabo_obj <- FindClusters(metabo_obj, resolution = 0.3)

plan("multisession", workers = thread)
metabo_obj <- FindClusters(metabo_obj, resolution = 0.4)

plan("multisession", workers = thread)
metabo_obj <- FindClusters(metabo_obj, resolution = 0.5)

plan("multisession", workers = thread)
metabo_obj <- FindClusters(metabo_obj, resolution = 0.6)

plan("multisession", workers = thread)
metabo_obj <- FindClusters(metabo_obj, resolution = 0.7)

plan("multisession", workers = thread)
metabo_obj <- FindClusters(metabo_obj, resolution = 0.8)

plan("multisession", workers = thread)
metabo_obj <- FindClusters(metabo_obj, resolution = 0.9)

plan("multisession", workers = thread)
metabo_obj <- FindClusters(metabo_obj, resolution = 1)

plan("multisession", workers = thread)
metabo_obj <- FindClusters(metabo_obj, resolution = 1.1)

plan("multisession", workers = thread)
metabo_obj <- FindClusters(metabo_obj, resolution = 1.2)

plan("multisession", workers = thread)
metabo_obj <- FindClusters(metabo_obj, resolution = 1.3)

plan("multisession", workers = thread)
metabo_obj <- FindClusters(metabo_obj, resolution = 1.4)
```

```{r}
clustree(metabo_obj, prefix = "RNA_snn_res.")
```
